Non-word repetition in Yélî Dnye

### 2020-09-16 manuscript.tex
bug fix

\begin{itemize}
\tightlist
\item
  look through this manuscript for bits of text to be written by \textbf{mc}
\item
  do you think Figure 2 should be done with phonetic symbols? If yes, could you try to implement that?
\item
  double check demo -- probably start with the xlsx version that has the children's first names, date of birth, and date of test. I looked again at my local files (still haven't checked the labs' ones) and I'm surprised I have maternal education for more children than I have birth order for\ldots{} If you correct this, to generate the csv I also round the age, assign age.rounded to reported age (in years) when DOB is missing, remove first names, DOB, DOTest and age.precise before pushing to github
\item
  double check my frequency entries in segments.xlsx using the search function in \url{https://phoible.org/parameters}; take a look at \url{http://phoible.github.io/conventions/} in case you see something about double articulation (I didn't find tp or lBj)
\item
  double check phonological implementation of stimuli in stimuli.xlsx
\item
  for both of the above, if you make any changes, they need to be re-saved as UTF16 txt file, and then converted (see remember above)
\item
  double check also stim\_seg\_freq -- why are some items still with NA freq?
\item
  draft discussion
\end{itemize}

@@

We first checked whether NWR scores varied between first and subsequent presentations of an item by averaging word-level scores at the participant level separately for first attempts and subsequent repetitions. We excluded 3 children who did not have data for one of these two types. As shown in Figure 1, participants' mean word-level scores became more heterogeneous in subsequent repetitions. Surprisingly, NWR scores for subsequent repetitions (M = 39, SD = 26)
were on average lower than first ones (M = 65, SD = 15), t(38) = 6.60, p = 0. Given the uncertainty in whether previous work used only the first or all repetitions, and since scores declined and became more heterogeneous in subsequent repetitions, the rest of the analyses focus on only the first repetitions.

@@

The overall NWR score was M = 61\% (SD = 15\%). Scores based on phonemes are even higher M = 77\% (SD = 10\%). The phoneme-based normalized Levenshtein distance was M = 23\% (SD = 10\%), meaning that about a fifth of phonemes were were substituted, inserted, or deleted. Notice that the normalized Levenshtein distance is the complement of phoneme-based scores.

@@

We could include 885 observations (\textbf{why -- shouldn't it be 1k??}), from 41 children producing in any given trial one of 41 potential target words. The analysis revealed a significant estimate for the average cross-linguistic frequency of phones in the target words (ß = 0.81, SE ß = 0.18, p = 0): Target words with phones found more frequently across languages had a higher proportion of words that were correctly repeated, as clear in Figure 2. Age was also a significant predictor (ß = 0.32, SE ß = 0.13, p = 0.01).

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/Fig2-xling_freq-1.pdf}
\caption{(\#fig:Fig2-xling\_freq)NWR scores for individual target words as a function of the average frequency with which each phone is found across languages.}
\end{figure}

We next checked whether this may actually be due to frequency of the sounds within the language. We estimated frequency of the phones present in the stimuli in a corpus of child-centered recordings ({\textbf{???}}), and applied the natural logarithm. As before, unattested sounds were not considered (i.e., they were declared NA so that they do not count for analyses). Phone frequency estimated from this corpus were correlated with cross-linguistic phone frequencies (r(27)=0.52, =0.00). Therefore, we fit another mixed logistic regression, this time declaring as fixed effects scaled cross-linguistic and corpus frequencies (averaged across all attested phones within each stimulus item), in addition to age. As before, the model contained random slopes for both child ID and target. In this model, both cross-linguistic frequency (ß = 0.81, SE ß = 0.27, p = 0.00) and age (ß = 0.32, SE ß = 0.13, p = 0.01) were significant predictors of word-level NWR scores, but corpus frequency (ß = 0.00, SE ß = 0.25, p = 1.00) was not.

@@

As for substitutions, it was as common for a nasal vowel to be produced as an oral vowel as vice versa (34 oral target vowels produced as nasal vowels, 31 nasal target vowels produced as oral vowels). Substitutions in which the oral nature was preserved but the quality of the vowel was changed were a great deal more common than changes in quality among nasal vowels (112 oral vowels produced with a different quality; 13 nasal vowels produced with a different quality). As for consonants, asymmetries were very marked with more complex consonants (specifically dptpkpkmknmbghlv) mispronounced as simple ones (specifically mnlwyvdgptkfhch, 117) than vice versa (2). Simple consonants were mispronounced as other simple consonants quite frequently (69 simple consonants mispronounced as other simple ones, compared to 33 complex ones).

@@

As for substitutions, it was as common for a nasal vowel to be produced as an oral vowel as vice versa (34 oral target vowels produced as nasal vowels or about 0.98\% out of all oral vowel targets, 31 nasal target vowels produced as oral vowels or about 2.37\% out of all nasal vowel targets). Substitutions in which the oral nature was preserved but the quality of the vowel was changed were a great deal more common than changes in quality among nasal vowels (112 oral vowels produced with a different quality or about 1.82\% out of all oral vowel targets; 13 nasal vowels produced with a different quality or about 2.04\% out of all nasal vowel targets). As for consonants, asymmetries were very marked with more complex consonants (specifically dptpkpkmknmbghlv) mispronounced as simple ones (specifically mnlwyvdgptkfhch, 117 times or about 1.95\% out of all complex consonant targets) than vice versa (2 times or about 1.95\% out of all simple consonant targets). Simple consonants were mispronounced as other simple consonants quite frequently (69 simple consonants mispronounced as other simple ones or about 0.38\% out of all simple consonant targets, compared to 33 complex ones or about 0.99\% out of all complex consonant targets).

@@

Our final exploratory analysis assessed whether variation in scores was structured by factors that vary across individuals. As shown in Figure 2, there was a greater deal of variance in earlier than later ages, with significantly higher NWR scores for older children (Spearman's rank correlation, given inequality of variance, rho (1,483.87) = 0.35, p = 0.09). In contrast, there was no clear association between NWR scores, on the one hand, and sex (t (0.13) = 21.37, p = 0.90), birth order (data missing for 17 children, 2,600.01) = -0.13, p = 0.54) or maternal education (data missing for 2 children, 2,045.68) = 0.11, p = 0.61)

>>>>

\begin{itemize}
\tightlist
\item
  look through this manuscript for bits of text to be written by \textbf{mc}
\item
  double check demo -- probably start with the xlsx version that has the children's first names, date of birth, and date of test. I looked again at my local files (still haven't checked the labs' ones) and I'm surprised I have maternal education for more children than I have birth order for\ldots{} If you correct this, to generate the csv I also round the age, assign age.rounded to reported age (in years) when DOB is missing, remove first names, DOB, DOTest and age.precise before pushing to github
\item
  double check my frequency entries in segments.xlsx using the search function in \url{https://phoible.org/parameters}; take a look at \url{http://phoible.github.io/conventions/} in case you see something about double articulation (I didn't find tp or lBj)
\item
  double check also stim\_seg\_freq -- why are some items still with NA freq?
\item
  draft discussion
\end{itemize}

@@

We first checked whether NWR scores varied between first and subsequent presentations of an item by averaging word-level scores at the participant level separately for first attempts and subsequent repetitions. We excluded 3 children who did not have data for one of these two types. As shown in Figure 1, participants' mean word-level scores became more heterogeneous in subsequent repetitions. Surprisingly, NWR scores for subsequent repetitions (M = 40, SD = 26)
were on average lower than first ones (M = 64, SD = 15), t(38) = 6.28, p = 0. Given the uncertainty in whether previous work used only the first or all repetitions, and since scores declined and became more heterogeneous in subsequent repetitions, the rest of the analyses focus on only the first repetitions.

@@

The overall NWR score was M = 60\% (SD = 15\%). Scores based on phonemes are even higher M = 77\% (SD = 10\%). The phoneme-based normalized Levenshtein distance was M = 23\% (SD = 10\%), meaning that about a fifth of phonemes were were substituted, inserted, or deleted. Notice that the normalized Levenshtein distance is the complement of phoneme-based scores.

@@

We could include 885 observations (\textbf{why -- shouldn't it be 1k??}), from 41 children producing in any given trial one of 41 potential target words. The analysis revealed a significant estimate for the average cross-linguistic frequency of phones in the target words (ß = 0.82, SE ß = 0.18, p = 0): Target words with phones found more frequently across languages had a higher proportion of words that were correctly repeated, as clear in Figure 2. Age was also a significant predictor (ß = 0.33, SE ß = 0.13, p = 0.01).

\begin{figure}[!t]

{\centering \includegraphics[width=0.65\linewidth]{nwr.by.freq.ITEM}

\caption{NWR scores for individual target words as a function of the average frequency with which each phone is found across languages.}\label{fig:Fig2-xling-freq}
\end{figure}

We next checked whether this may actually be due to frequency of the sounds within the language. We estimated frequency of the phones present in the stimuli in a corpus of child-centered recordings (Casillas et al., 2020), and applied the natural logarithm. As before, unattested sounds were not considered (i.e., they were declared NA so that they do not count for analyses). Phone frequency estimated from this corpus were correlated with cross-linguistic phone frequencies (r(27)=0.52, =0.00). Therefore, we fit another mixed logistic regression, this time declaring as fixed effects scaled cross-linguistic and corpus frequencies (averaged across all attested phones within each stimulus item), in addition to age. As before, the model contained random slopes for both child ID and target. In this model, both cross-linguistic frequency (ß = 0.82, SE ß = 0.27, p = 0.00) and age (ß = 0.33, SE ß = 0.13, p = 0.01) were significant predictors of word-level NWR scores, but corpus frequency (ß = 0, SE ß = 0.25, p = 1.00) was not.

@@

As for substitutions, it was as common for a nasal vowel to be produced as an oral vowel as vice versa (34 oral target vowels produced as nasal vowels, 30 nasal target vowels produced as oral vowels). Substitutions in which the oral nature was preserved but the quality of the vowel was changed were a great deal more common than changes in quality among nasal vowels (110 oral vowels produced with a different quality; 13 nasal vowels produced with a different quality). As for consonants, asymmetries were very marked with more complex consonants (specifically dptpkpkmknmbghlv) mispronounced as simple ones (specifically mnlwyvdgptkfhch, 115) than vice versa (2). Simple consonants were mispronounced as other simple consonants quite frequently (69 simple consonants mispronounced as other simple ones, compared to 33 complex ones).

@@

As for substitutions, it was as common for a nasal vowel to be produced as an oral vowel as vice versa (34 oral target vowels produced as nasal vowels or about 0.99\% out of all oral vowel targets, 30 nasal target vowels produced as oral vowels or about 2.37\% out of all nasal vowel targets). Substitutions in which the oral nature was preserved but the quality of the vowel was changed were a great deal more common than changes in quality among nasal vowels (110 oral vowels produced with a different quality or about 1.86\% out of all oral vowel targets; 13 nasal vowels produced with a different quality or about 2.06\% out of all nasal vowel targets). As for consonants, asymmetries were very marked with more complex consonants (specifically dptpkpkmknmbghlv) mispronounced as simple ones (specifically mnlwyvdgptkfhch, 115 times or about 1.96\% out of all complex consonant targets) than vice versa (2 times or about 1.96\% out of all simple consonant targets). Simple consonants were mispronounced as other simple consonants quite frequently (69 simple consonants mispronounced as other simple ones or about 0.39\% out of all simple consonant targets, compared to 33 complex ones or about 1.02\% out of all complex consonant targets).

@@

Our final exploratory analysis assessed whether variation in scores was structured by factors that vary across individuals. As shown in Figure 2, there was a greater deal of variance in earlier than later ages, with significantly higher NWR scores for older children (Spearman's rank correlation, given inequality of variance, rho (1,421.71) = 0.38, p = 0.07). In contrast, there was no clear association between NWR scores, on the one hand, and sex (t (0.03) = 21.40, p = 0.98), birth order (data missing for 17 children, 2,652.32) = -0.15, p = 0.47) or maternal education (data missing for 2 children, 2,101.60) = 0.09, p = 0.69)
